### 2018-10-09 post-linearity-2018/LAM.tex
Strong normalization.

\textbf{Additional Results.} In addition to the results given so far, we
also proved that the sequent calculus formalization is equivalent to the
natural deduction formalization given here
(Appendix~\ref{app:seq-nat-equiv}). Furthermore, we proved strong
normalization of CNC logic via a translation to LNL logic
(Appendix~\ref{app:strong-normalization}).

>>>>

We also proved that the sequent calculus formalization given in
Figure~\ref{fig:CNC-sequent-calculus} is equivalent to the typing rules (or
else called the natural deduction formalization) given in
Figure~\ref{fig:CNC-typing-rules} are equivalent, as stated in the following
theorem.
\begin{theorem}
  \label{thm:sc-nd-equiv}
  The sequent calculus ($\mathit{SC}$) and natural deduction ($\mathit{ND}$)
  formalizations for CNC logic are equivalent in the sense that there are
  two mappins $N:\mathit{SC}\rightarrow\mathit{ND}$ and
  $S:\mathit{ND}\rightarrow\mathit{SC}$ that map each rule in $\mathit{SC}$
  to a proof in $\mathit{ND}$, and each rule in $\mathit{ND}$ to a proof
  in $\mathit{SC}$, respectively.
\end{theorem}
\begin{proof}
  The proof is done case by case on each rule in the sequence calculus and
  natural deduction formalizations. It is obvious that the axioms in one
  formalization can be mapped to the axioms in the other. The introduction
  rules in $\mathit{ND}$ are mapped to the right rules in $\mathit{SC}$, and
  vice versa. The elimination rules and lefts rules are mapped to each other
  with some fiddling. For instance, the elimination rule for the
  non-commutative tensor is mapped to the following proof in $\mathit{SC}$:
  \begin{center}
    \scriptsize
    \begin{math}
      $$\mprset{flushleft}
      \inferrule* [right={\scriptsize $\ElledruleTXXcutName$}] {
        {[[I |-c t1 : X (*) Y]]} \\
        $$\mprset{flushleft}
        \inferrule* [right={\scriptsize $\ElledruleTXXtenLName$}] {
          {[[P1, x : X, y : Y, P2 |-c t2 : Z]]}
        }{[[P1, z : X (*) Y, P2 |-c let z : X (*) Y be x (*) y in t2 : Z]]}
      }{[[P1, I, P2 |-c [t1 / z](let z : X (*) Y be x (*) y in t2) : Z]]}
    \end{math}
  \end{center}
  The full proof is in Appendix~\ref{app:sc-nd-equiv}.
\end{proof}

Finally, we proved strong normalization of CNC logic via a translation to
LNL logic. The proof is in Appendix~\ref{app:strong-normalization}. We first
define a function $CL$ that maps our term assignment for natural deduction
to Benton's term assignment, and then shows that if there is reduction
$[[t1 ~> t2]]$ (resp. $[[s1 ~> s2]]$) in CNC, then there is a reduction
$CL(t_1)\leadsto CL(t_2)$ (resp. $CL(s_1)\leadsto CL(s_2)$) in LNL. Recall
that a LNL model consists of an adjunction $F:\cat{C}\dashv\cat{L}:G$ in
which $\cat{C}$ is a cartesian closed category and $\cat{L}$ is a SMCC.
The basic idea of mapping a LAM model to a LNL model is mapping the SMCC in
LAM to the cartesian closed category in LNL, and the Lambek category in LAM
to the SMCC in LNL. For instance, $s_1\triangleright s_2$ in CNC, where
$[[s1]]$ and $[[s2]]$ are atomic terms, maps to $s_1\otimes s_2$ in LNL.

### 2018-10-09 post-linearity-2018/adjoint-models.tex
Strong normalization.

\textbf{An isomorphism}. Let $(\cat{C},\cat{L},F,G,\eta,\varepsilon)$
be a LAM, where $(F,\m{})$ and $(G,\n{})$ are monoidal
functors. Similarly as in Benton's LNL model, $\m{X,Y} : FX \tri FY
\mto F(X \otimes Y)$ are components of a natural isomorphism, and
$\m{I} : I' \mto FI$ is an isomorphism. This is essential for modeling
certain rules of CNC logic, such as tensor elimination in natural
deduction.  We define the inverses of $\m{X,Y}:FX\tri FY\rightarrow
F(X\otimes Y)$ and $\m{I}:I'\rightarrow FI$ as:
\begin{mathpar}
\bfig
  \morphism<1000,0>[\p{X,Y}:F(X\otimes Y)`F(GFX\otimes GFY);F(\eta_X\otimes\eta_Y)]
  \morphism(1000,0)<900,0>[F(GFX\otimes GFY)`FG(FX\tri FY);F\n{FX,FY}]
  \morphism(1900,0)<750,0>[FG(FX\tri FY)`FX\tri FY;\varepsilon_{FX\tri FX}]
\efig
\and
\bfig
  \morphism<500,0>[\p{I}:FI`FGI';F\n{I'}]
  \morphism(500,0)<400,0>[FGI'`I';\varepsilon_{I'}]
\efig
\end{mathpar}
We can now see that this straightforwardly follows from Benton's proof.

\begin{theorem}
\label{thm:m-natural-iso}
  $\m{X,Y}$ are components of a natural isomorphism and their inverses are $\p{X,Y}$. 
\end{theorem}
\begin{proof}
  This proof follows by a diagram chase using the adjunction in the
  definition of a LAM and naturality. The detail of the proof is in
  Appendix~\ref{app:m-natural-iso}.
\end{proof}

\begin{theorem}
\label{thm:m-iso}
  $\m{I}$ is an isomorphism and its inverse is $\p{I}$.
\end{theorem}
\begin{proof}
  This is equivalent to equations $\m{I}\circ\p{I}=id_{FI}$ and
  $\p{I}\circ\m{I}=id_{I'}$, equivalent to the following diagrams, which
  commute because $\varepsilon$ is a monoidal natural transformation.
  \begin{mathpar}
  \bfig
    \square/->`=`->`<-/<400,400>[FI`FGI'`FI`I';F\n{I'}``\varepsilon_{I'}`\m{I}]
  \efig
  \and
  \bfig
    \square/->`=`->`<-/<400,400>[I'`FI`I'`FGI';\m{I}``F\n{I'}`\varepsilon_{I'}]
  \efig
  \end{mathpar}
\end{proof}

>>>>

\textbf{An isomorphism}. Let $(\cat{C},\cat{L},F,G,\eta,\varepsilon)$
be a LAM, where $(F,\m{})$ and $(G,\n{})$ are monoidal
functors. Similarly as in Benton's LNL model, $\m{X,Y} : FX \tri FY
\mto F(X \otimes Y)$ are components of a natural isomorphism, and
$\m{I} : I' \mto FI$ is an isomorphism. This is essential for modeling
certain rules of CNC logic, such as tensor elimination in natural
deduction.  We define the inverses of $\m{X,Y}:FX\tri FY\rightarrow
F(X\otimes Y)$ and $\m{I}:I'\rightarrow FI$ as:
\vspace{-1em}
\begin{mathpar}
\footnotesize
\bfig
  \morphism<1000,0>[\p{X,Y}:F(X\otimes Y)`F(GFX\otimes GFY);F(\eta_X\otimes\eta_Y)]
  \morphism(1000,0)<900,0>[F(GFX\otimes GFY)`FG(FX\tri FY);F\n{FX,FY}]
  \morphism(1900,0)<750,0>[FG(FX\tri FY)`FX\tri FY;\varepsilon_{FX\tri FX}]
\efig
\end{mathpar}
\vspace{-1.6em}
\begin{mathpar}
\footnotesize
\bfig
  \morphism<500,0>[\p{I}:FI`FGI';F\n{I'}]
  \morphism(500,0)<400,0>[FGI'`I';\varepsilon_{I'}]
\efig
\end{mathpar}
\begin{theorem}
\label{thm:m-iso}
  $\m{I}$ is an isomorphism and its inverse is $\p{I}$.
\end{theorem}
\begin{proof}
  This is equivalent to equations $\m{I}\circ\p{I}=id_{FI}$ and
  $\p{I}\circ\m{I}=id_{I'}$, equivalent to the following diagrams, which
  commute because $\varepsilon$ is a monoidal natural transformation.
  \begin{mathpar}
  \footnotesize
  \bfig
    \square/->`=`->`<-/<400,400>[FI`FGI'`FI`I';F\n{I'}``\varepsilon_{I'}`\m{I}]
  \efig
  \and
  \bfig
    \square/->`=`->`<-/<400,400>[I'`FI`I'`FGI';\m{I}``F\n{I'}`\varepsilon_{I'}]
  \efig
  \end{mathpar}
\end{proof}

\begin{theorem}
\label{thm:m-natural-iso}
  $\m{X,Y}$ are components of a natural isomorphism and their inverses are $\p{X,Y}$. 
\end{theorem}
\begin{proof}
  This is equivalent to that the two equations
  $\m{X,Y}\circ\p{X,Y}=id_{F(X\otimes Y)}$ and
  $\p{X,Y}\circ\m{X,Y}=id_{FX\tri FX}$ hold, which is proved because the
  following diagrams commute: (1)-adjunction; (2)-$\eta$ is a monoidal
  natural transformation; (3)-naturality of $\varepsilon$; (4)-adjunction;
  (5)-naturality of $\m{}$; (6)-$\varepsilon$ is a monoidal natural
  transformation.
  \begin{mathpar}
  \footnotesize
  \bfig
    \square/->`=`->`/<1600,800>[
      F(X\otimes Y)`F(GFX\otimes GFY)`F(X\otimes Y)`FG(FX\tri FY);
      F(\eta_X\otimes\eta_Y)``F\n{FX,FY}`]
    \Atrianglepair|mmmbb|/->``->`<-`<-/<800,400>[
      FGF(X\otimes Y)`F(X\otimes Y)`FX\tri FY`FG(FX\tri FY);
      \varepsilon_{F(X\otimes Y)}``FG\m{X,Y}`\m{X,Y}`\varepsilon_{FX\tri FY}]
    \morphism(0,800)|m|<800,-400>[F(X\otimes Y)`FGF(X\otimes Y);F\eta_{X\otimes Y}]
    \morphism(300,450)//<0,0>[`;(1)]
    \morphism(1200,600)//<0,0>[`;(2)]
    \morphism(800,200)//<0,0>[`;(3)]
  \efig
  \end{mathpar}
  \begin{mathpar}
  \footnotesize
  \bfig
    \square/->`=`->`/<1800,800>[
      FX\tri FY`F(X\otimes Y)`FX\tri FY`F(GFX\otimes GFY);
      \m{X,Y}``F(\eta_X\otimes\eta_Y)`]
    \Atrianglepair|mmmbb|/->``->`<-`<-/<900,400>[
      FGFX\tri FGFY`FX\tri FY`FG(FX\tri FY)`F(GFX\otimes GFY);
      \varepsilon_{FX}\tri\varepsilon_{FY}``\m{GFX,GFY}`\varepsilon_{FX\tri FY}`F\n{FX,FY}]
    \morphism(0,800)|m|<900,-400>[FX\tri FY`FGFX\tri FGFY;F\eta_X\tri F\eta_Y]
    \morphism(300,450)//<0,0>[`;(4)]
    \morphism(1300,600)//<0,0>[`;(5)]
    \morphism(900,200)//<0,0>[`;(6)]
  \efig
  \end{mathpar}
\end{proof}
